---
title: "Statistical analysis"
author: "Randi Maud Mork"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

### Loading packages

```{r}
library(fpc)
library(dbscan)
library(factoextra)
library(dplyr)
library(glmnet)
library(tidyr)
```

# Observed data

## Data preprocessing

```{r}

# standardizing
Conv_ind_soundings_sd <- as.data.frame(scale(drop_na(Convective_indices_soundings[,-c(1,5)])))%>%
  mutate(conv = factor(Convective_indices_soundings_classified$conv, levels = c(0,1)))
summary(Conv_ind_soundings_sd)

# log-transforming
Conv_ind_soundings_log <- Convective_indices_soundings%>%
  mutate(CTP_log = log(CTP), HI_low_log = log(HI_low), 
         SM_Morning_log = log(SM_Morning), BR_Morning_log = log(BR_Morning),
         conv = Convective_indices_soundings_classified$conv)
summary(Conv_ind_soundings_log)

```

## Clustering, DBSCAN

```{r}
dbscan::kNNdistplot(Conv_ind_soundings_sd[,-c(3,8)], k =  3)
abline(h = 1.5, lty = 2)
cluster.mod <- fpc::dbscan(Conv_ind_soundings_sd[, -c(3,8)] , eps = 1.5, MinPts = 3, scale = , method = "raw")

fviz_cluster(cluster.mod, Conv_ind_soundings_sd[,-c(3,8)], stand = F, ellipse = T,geom = "point", choose.vars = c("CTP", "HI_low"))
plot(cluster.mod, Conv_ind_soundings_sd[, -5], main = "DBSCAN")


CTP_HI_low_clustered <- cbind(Convective_indices_soundings_classified,cluster.mod$cluster)%>%
  mutate(cluster=factor(`cluster.mod$cluster`))
```

## Generalised linear model

```{r}
# Standardized data
glm.obj <- glm(conv ~ ., data = Conv_ind_soundings_sd, family = "binomial")
plot(glm.obj, which = c(1,2), caption = c("Resid vs fit,FUll dataset, standardized", "QQ-plot,Full dataset sd"))
summary(glm.obj)

# reduced variables
glm.red.obj <- glm(conv ~ CTP+HI_low+SM_Morning, data = Conv_ind_soundings_sd, family = "binomial")
plot(glm.red.obj, which = c(1,2), caption = c("Resid vs fit,reduced dataset, standardized", "QQ-plot, reduced dataset sd"))
summary(glm.red.obj)
Conv_prob <- predict(glm.red.obj, newdata = Conv_ind_soundings_sd, type = "response")
Conv_prob


## log transform to see for better fit of residual and QQ - plots
glm.log.obj <- glm(conv ~ CTP_log+HI_low_log+SM_Morning_log, data = Conv_ind_soundings_log, family = "binomial")
plot(glm.log.obj, which = c(1,2), caption = c("Resid vs fit, reduced dataset, log", "QQ-plot, reduced dataset, log"))
summary(glm.log.obj)


```
In the analysis of the observed data the model with the standardised dataset fits the data best and this model is chosen for further analysis.


# Reanalysis 

## Data preprocessing

```{r}
#standardization and variable selection
CTP_HI_low_reanalysis_model <- as.data.frame(scale(Reanalysis_classified[,c(2,3,4,7,8,9)])) %>%
  mutate(Precip_type = Reanalysis_classified$Precip_type, Convection = Reanalysis_classified$Convection)

# log-transformation
CTP_HI_low_reanalysis_model_log <- as.data.frame(log(Reanalysis_classified[,c(2,3,4,7,8,9)])) %>%
  mutate(Precip_type = Reanalysis_classified$Precip_type, Convection = Reanalysis_classified$Convection)%>%drop_na()
summary(CTP_HI_low_reanalysis_model_log)


```

## Clustering DBSCAN

```{r}
dbscan::kNNdistplot(CTP_HI_low_reanalysis_model[,c(1,2,3,4)], k =  3)
abline(h = 0.9, lty = 2)
cluster.mod <- fpc::dbscan(CTP_HI_low_reanalysis_model[,c(1,2,3,4)],eps = 0.9, MinPts = 2)

fviz_cluster(cluster.mod, CTP_HI_low_reanalysis_model, stand = F, ellipse = T,geom = "point", choose.vars = c("CTP", "HI_low"))
plot(cluster.mod, CTP_HI_low_reanalysis_model[,c(1,2,3,4)], main = "DBSCAN")

CTP_HI_low_reanalysis_clustered <- cbind(Reanalysis_classified,cluster.mod$cluster)%>%
  mutate(cluster=factor(`cluster.mod$cluster`))
```


## Generalised linear model

```{r}
# Standardized values 
glm_Reanalysis <- glm(Convection ~ ., data = CTP_HI_low_reanalysis_model[,c(1,2,3,8)], family = "binomial")

summary(glm_Reanalysis)
plot(glm_Reanalysis, which = c(1,2))

 # log-transformed values 
glm_re_log <- glm(Convection ~ ., data = CTP_HI_low_reanalysis_model_log[,c(1,2,3,8)], family = "binomial")

summary(glm_re_log)
plot(glm_re_log, which = c(1,2))
Conv_prob <- predict(glm_re_log, newdata = CTP_HI_low_reanalysis_model_log[,c(1,2,3,8)], type = "response")

Conv_prob
```


In reanalysis data the log transformed data shows a model better fit to the data. The model with log-transformed data is therefore chosen. 


















